---
title: "MAGICAL visualization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MAGICAL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=T, results='hide', warning=FALSE, message=FALSE}
library(magical)
```

# MAGICAL visualzation

MAGICAL offers several ways of visualizing circuits, all of which make use of the package `karyoploteR` and `regioneR`.


## Read the results

Let's start by reading the results from the MAGICAL analysis, "MAGICAL_selected_regulatory_circuits.txt". We suggest running this demo with our [provided file](https://github.com/ZxZhou4150/magical/blob/R-package/tutorial/MAGICAL_selected_regulatory_circuits.txt), because as mentioned before, running MAGICAL can return very similar but slightly different results each time.

```{r}
# read the results
data = read.table('MAGICAL_selected_regulatory_circuits.txt', header = T, sep = "\t")
```

## Visualize the circuits

MAGICAL offers 3 ways of visualizing the circuits: by circuit index, gene, or peak. For either of these ways, MAGICAL will plot a certain region of one chromosome which includes the circuit(s). Basically, the following information will be manifested on the plot:

  * The chromosome id and coverage of the region
  * The *starting site* of the gene
  * The starting site and ending site of the peak
  * The top 3 (or less if there are less than 3) TFs that are most likely to bind to the peak.
  
Also, see the [advanced option](#advanced-option) section for an advanced visualization option.
  

### visualize by circuit index

With function `plot_circuits_with_idx`, you can visualize one circuit by its index.

```{r, fig.width=7, fig.height=4}
# visualize the first circuit
plot_circuits_with_idx(data, idx = 1)
```

### visualize by gene

With function `plot_circuits_with_gene`, you can visualize all circuits that contain a specific gene.

```{r, fig.width=7, fig.height=4}
# visualize circuits with gene ACOT11
plot_circuits_with_gene(data, gene = "ACOT11")
```


### visualize by peak

With function `plot_circuits_with_peak`, you can visualize all circuits that contain a specific peak.

```{r, fig.width=7, fig.height=4}
# visualize circuits with peak chr4:184817550-184818412
plot_circuits_with_peak(data, peak_chr = "chr4", peak_start = 184817550, peak_end = 184818412)
```


## Advanced option

All the 3 functions offer an advanced option, which allows you to see all the genes around the circuit region. This is useful if you want to see the specificity of the circuit(s) in the context of the whole region, which is, the other genes in that region are not involved in the circuit(s). 

This can be done by setting the parameter `gene_track` to `TRUE` (default is `FALSE`). After that, you need another parameter, `TxDb`, to specify the reference genome from which the function will query genes. Before that, you must download the corresponding TXDb package from Bioconductor. For example, if your reference genome is hg38, you need to download the package `TxDb.Hsapiens.UCSC.hg38.knownGene`.


```{r, fig.width=7, fig.height=4, warning=FALSE, message=FALSE}
## Here our reference genome is hg38

# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("TxDb.Hsapiens.UCSC.hg38.knownGene")

library(TxDb.Hsapiens.UCSC.hg38.knownGene)

# visualize the first circuit with gene track
plot_circuits_with_idx(data, 1 , T, TxDb.Hsapiens.UCSC.hg38.knownGene)
```
